\addcontentsline{toc}{chapter}{Literaturverzeichnis}

\begin{thebibliography}{99}

\bibitem{PostScript}
Adobe Systems Incorporated, 1990. \textsl{PostScript Language Reference Manual (second edition)}
Addison-Wesley, Reading, MA.

\bibitem{Assad92}
Assad, A.M.\ und Packard, N.H. 1992. "`Emergent Colonization in an Artificial Ecology."'
In \textsl{Toward a Practice of Autonomous Systems}, herausgegeben von F.J.\ Varela
und P. Bourgine, 143-152. MIT Press, Cambridge, MA.

\bibitem{Baas94}
Baas, N.A. 1994. "`Emergence, Hierarchies, and Hyperstructures."' In
\textsl{Artificial Life III}, herausgegeben von C.G. Langton, 515-537.
Addison-Wesley, Redwood City, CA.

% \bibitem{Batali94} Batali, J. 1994. "`Innate Biases and Critical Periods:
% Combining Evolution and Learning in the Acquisition of Syntax."'
% In \textsl{Artificial Life IV}, herausgegeben von R.A.\ Brooks und
% P.\ Maes, 160-171. MIT Press, Cambridge, Massachusetts.

% \bibitem{BedauPackard92}
% Bedau, M.A. und Packard, N.H. 1992. "`Measurement of Evolutionary Activity, Teleology, 
% and Life."' In \textsl{Artificial Life II}, herausgegeben von C.G.\ Langton 
% C.\ Taylor, J.D.\ Farmer und S.\ Rasmussen, 431-462. Santa Fe
% Institute Studies in the Sciences of Complexity, Proceedings Volume X.
% Addison-Wesley, Redwood City, CA.

\bibitem{deBoer92}
de Boer, M.J.M., Fracchia, D.\ und Prusinkiewicz, P. 1992. "`Analysis and Simulation of the Development of Cellular Layers."'
In \textsl{Artificial Life II}, herausgegeben von C.G.\ Langton 
C.\ Taylor, J.D.\ Farmer und S.\ Rasmussen, 465-483. Santa Fe
Institute Studies in the Sciences of Complexity, Proceedings Volume X.
Addison-Wesley, Redwood City, CA.

\bibitem{Boerlijst92}
Boerlijst, M.\ und Hogeweg, P. 1992. "`Self-Structuring and Selection: Spiral
Waves as a substrate for Prebiotic Evolution."'
In \textsl{Artificial Life II}, herausgegeben von C.G.\ Langton 
C.\ Taylor, J.D.\ Farmer und S.\ Rasmussen, 255-276. Santa Fe
Institute Studies in the Sciences of Complexity, Proceedings Volume X.
Addison-Wesley, Redwood City, CA.

\bibitem{Bonner88}
Bonner, J.T. 1988. \textsl{The Evolution of Complexity by Means of Natural Selection.}
Princeton University Press, Princeton, NJ.

\bibitem{CampbellLNCS88}
Campbell, D. 1989. "`Introduction to Nonlinear Phenomena."' In \textsl{Lectures in the Sciences of Complexity,}
herausgegeben von D. Stein, 3-103. Santa Fe
Institute Studies in the Sciences of Complexity. Addison-Wesley, Redwood City, CA.

\bibitem{Cariani92}
Cariani, P. 1992. "`Emergence and Artificial Life."'
In \textsl{Artificial Life II}, herausgegeben von C.G.\ Langton 
C.\ Taylor, J.D.\ Farmer und S.\ Rasmussen, 775-797. Santa Fe
Institute Studies in the Sciences of Complexity, Proceedings Volume X.
Addison-Wesley, Redwood City, CA.

\bibitem{Carroll95}
Carroll, S.B. 1995. "`Homeotic Genes and the Evolution of Arthropods and Chordates."'
\textsl{Nature} \textbf{376:} 479-485.

% \bibitem{Conrad89}
% Conrad, M.\ und Ritzki, M.M. 1989. "`The Artificial Worlds Approach to Emergent
% Evolution."' \textsl{BioSystems} \textbf{23:} 247-260.

\bibitem{Darwin}
Darwin, C. 1859. \textsl{On the Origin of Species by Means of natural Selection.} London, 1859.

\bibitem{Doyle94}
Doyle, J.J. 1994. "`Evolution of a Plant Homeotic Multigene Family: Toward Connecting Molecular Systematics
and Molecular Developmental Genetics."' \textsl{Syst. Biol.} \textbf{43:} 307-328.

\bibitem{Eigen71}
Eigen, M. 1971. "`Selforganization of Matter and the Evolution of Biological
Macromolecules."' \textsl{Naturwissenschaften} \textbf{58:} 465-526.

\bibitem{Felsenstein88}
Felsenstein, J. 1988. "`Phylogenies from Molecular Sequences: Inference and Reliability."'
\textsl{Annu.\ Rev.\ Genet.} \textbf{22:} 521-565.

\bibitem{PHYLIP}
Felsenstein, J. 1989. "`PHYLIP -- Phylogeny Inference Package (Version 3.2)"'
\textsl{Cladistics} \textbf{5:} 164-166.

\bibitem{Forst95}
Forst, C.V., Reidys C.\ und Weber, J. 1995. "`Evolutionary Dynamics and Optimization."'
In \textsl{Advances in Artificial Life}, herausgegeben von F.\ Mor{\'a}n, A.\ Moreno,
J.J.\ Merelo und P.\ Chac{\'o}n, 128-147. Springer Verlag, Berlin Heidelberg.

\bibitem{Gardner70}
Gardner, M. 1970. "`Mathematical Games: The Fantastic Combinations of John Conway's New
Solitaire Game 'Life'."' \textsl{Scientific American} 4/1970: 120-123.

% \bibitem{Gleick87}
% Gleick, J.\ 1987. \textsl{Chaos.} Penguin Books, New York.

\bibitem{GCG}
Genetics Computer Group, 1994. \textsl{Program Manual for the Wisconsin Package,
            Version 8.} September 1994, Genetics Computer Group,
             575 Science Drive, Madison, Wisconsin, USA  53711

\bibitem{Gerhardt89}
Gerhardt, M.\ und Schuster, H. 1989. "`A Cellular Automaton Describing the Formation of Spatially
Ordered Structures in Chemical Systems."' \textsl{Physica D} \textbf{36:} 209-221.

\bibitem{Goldberg89}
Goldberg, D.E. 1989. \textsl{Genetic Algorithms in Search, Optimization and Machine Learning.}
Addison-Wesley, New York, NY.

\bibitem{Gutowitz95}
Gutowitz, H.\ und Langton, C.G. 1995. "`Mean Field Theory of the Edge of Chaos."'
In \textsl{Advances in Artificial Life}, herausgegeben von F.\ Mor{\'a}n, A.\ Moreno,
J.J.\ Merelo und P.\ Chac{\'o}n, 53-64. Springer Verlag, Berlin Heidelberg.

\bibitem{Hainzl}
Hainzl, J. 1981. \textsl{Mathematik f"ur Naturwissenschaftler.} Teubner Studienb"ucher
Mathematik, B.G.\ Teubner, Stuttgart.

\bibitem{Hamilton93}
Hamilton, P. 1993. \textsl{Eine Sprache zur Simulation der Morphologie von Dendriten.} Dissertation
an der Universit"at Bonn, Abt.\ Theoretische Biologie.

\bibitem{Harbison91}
Harbison, S.P.\ und Steele, G.L. 1991. \textsl{C. A Reference Manual.}
Prentice Hall, Englewood Cliffs, NJ.

\bibitem{Higgs92}
Higgs. P.G.\ und Derrida, P. 1992. "`Genetic Distance and Species Formation
in Evolving Populations."' \textsl{J.Mol.Evol.} \textbf{35:} 454-465.

% \bibitem{Horgan95}
% Horgan, J. 1995. "`From Complexity to Perplexity."' \textsl{Scientific American} 6/1995: 74-79.

\bibitem{Horn92}
Horn, J. 1992. "`Measuring the Evolving Complexity of Stimulus-Response Organisms."'
In \textsl{Towards a Practice of Autonomous Systems}, herausgegeben von
F.J.\ Varela und P.\ Bourgine, 365-374. MIT Press, Cambridge, Massachusetts.

\bibitem{Joyce92}
Joyce, G.F. 1992. "`Directed Molecular Evolution." \textsl{Scientific American} 12/1992: 48-55.

\bibitem{Kauffman92} Kauffman, S.\ und Johnsen, S. 1992. "`Co-Evolution to the Edge of Chaos:
Coupled Fitness Landscape, Poised States, and Co-Evolutionary Avalanches."'
In \textsl{Artificial Life II}, herausgegeben von C.G.\ Langton 
C.\ Taylor, J.D.\ Farmer und S.\ Rasmussen, 325-369. Santa Fe
Institute Studies in the Sciences of Complexity, Proceedings Volume X.
Addison-Wesley, Redwood City, CA.

\bibitem{Keller94}
Keller, U. 1994. \textsl{Modelltheoretische Untersuchung evolutionsbiologischer Mechanismen
mit Hilfe Genetischer Algorithmen.} Inaugural-Dissertation zur Erlangung des
Doktorgrades der Mathematisch-Naturwissenschaftlichen Fakult"at der Universit"at
zu K"oln.

\bibitem{Kim95}
Kim, J.T. 1995. "`Distanzverteilungskomplexit"at: Ein Ma"s f"ur die strukturierte Diversit"at
evolvierender Populationen."' In \textsl{Proceedings des Workshops "`Artificial Life"'},
herausgegeben von K.\ Dautenhahn, T.\ Christaller, J.\ Heistermann, R.\ Hofest"adt, J.\ Kaiser,
H.-G.\ Lipinski, C.\ M"uller-Schloer, W.\ Schiffmann und D.\ Sch"utt. GMD-Studien, Nr.\ 271, Sankt Augustin.

\bibitem{Kim95a}
Kim, J.T. 1995. "`Using Distance Distributions to Measure Complexity of Populations."'
In \textsl{1993 Lectures in Complex Systems,} herausgegeben von D.L.\ Stein und L.\ Nadel.
Santa Fe Institute Studies in the Sciences of Complexity, Lectures Volume VI.
Addison-Wesley, Redwood City, CA.

\bibitem{Kornberg80}
Kornberg, A. 1980. \textsl{DNA Replication.} W.H.\ Freeman and Company, San Francisco.

\bibitem{KuhnerFelsenstein}
Kuhner, M.K. and Felsenstein, J. 1994. "`A Simulation Comparison of Phylogeny Algorithms
under Equal and Unequal Evolutionary Rates."' \textsl{Mol. Biol. Evol.} 11(3): 459-468.

\bibitem{ALifeI}
Langton, C.G. 1989. \textsl{Artificial Life}
Santa Fe Institute Studies in the Sciences of 
Complexity, Proceedings Volume VI. Addison-Wesley, Redwood City, CA.

\bibitem{ALifeII}
Langton, C.G., Taylor, C., Farmer, J.D.\ und Rasmussen, S. 1992.
\textsl{Artificial Life II.} Santa Fe
Institute Studies in the Sciences of Complexity, Proceedings Volume X.
Addison-Wesley, Redwood City, CA.

\bibitem{Langton90}
Langton, C.G. 1990. "`Computation at the Edge of Chaos: Phase Transitions and Emergent Computation."'
\textsl{Physica D} \textbf{42:} 12-37.

\bibitem{Langton92}
Langton, C.G. 1992. "`Life at the Edge of Chaos."'
In \textsl{Artificial Life II}, herausgegeben von C.G.\ Langton 
C.\ Taylor, J.D.\ Farmer und S.\ Rasmussen, 41-91. Santa Fe
Institute Studies in the Sciences of Complexity, Proceedings Volume X.
Addison-Wesley, Redwood City, CA.

\bibitem{Lindgren92}
Lindgren, K. 1992. "`Evolutionary Phenomena in Simple Dynamics."'
In \textsl{Artificial Life II}, herausgegeben von C.G.\ Langton 
C.\ Taylor, J.D.\ Farmer und S.\ Rasmussen, 295-312. Santa Fe
Institute Studies in the Sciences of Complexity, Proceedings Volume X.
Addison-Wesley, Redwood City, CA.

\bibitem{May95}
May, R.M., Bohoeffer, S.\ und Nowak, M.A. 1995. "`Spatial Games and Evolution of Cooperation."'
In \textsl{Advances in Artificial Life}, herausgegeben von F.\ Mor{\'a}n, A.\ Moreno,
J.J.\ Merelo und P.\ Chac{\'o}n, 749-759. Springer Verlag, Berlin Heidelberg.

\bibitem{Maynard89}
Maynard Smith, J. 1989. \textsl{Evolutionary Genetics.} Oxford University Press, Oxford.

\bibitem{Maley94}
Maley, C. 1994. "`The Coevolution of Mutation Rates."'
In \textsl{Artificial Life IV}, herausgegeben von R.A.\ Brooks und
P.\ Maes, 219-233. MIT Press, Cambridge, Massachusetts.

% \bibitem{Meissner95}
% Mei"sner, R.\ und Theres, K. 1995. "`Iolation and Characterization of the
% Tomato Homeobox Gene THOM1."' \textsl{Planta} \textbf{195:} 541-547.

\bibitem{NeedlemanWunsch}
Needleman, S.B.\ und Wunsch, C.D. 1970. "`A General Method Applicable to the Search for
Similarities in the Amino Acid Sequence of Two Proteins."' \textsl{J.\ Mol.\ Biol.} \textbf{48:} 443-453.

\bibitem{Oberschelp92}
Oberschelp, W.\ und Vossen, G. 1992. \textsl{Rechneraufbau und Rechnerstrukturen.}
R.\ Oldenbourg Verlag, M"unchen Wien.

\bibitem{Packard89}
Packard, N.H. 1989. "`Intrinsic Adaptation in a Simple Model for Evolution."' In \textsl{Artificial 
Life}, herausgegeben von C.G. Langton, 141-155. Santa Fe Institute Studies in the Sciences 
of Complexity, Proceedings Volume VI. Addison-Wesley, Redwood City, CA.

% \bibitem{Parisi92} Parisi, D., Nolfi, S.\ und Cecconi, F. 1992.
% "`Learning, Behavior, and Evolution."'
% In \textsl{Towards a Practice of Autonomous Systems}, herausgegeben von
% F.J.\ Varela und P.\ Bourgine, 207-216. MIT Press, Cambridge, Massachusetts.

\bibitem{Pellegrini95}
Pellegrini, L., Tan, S.\ und Richmond, T.J. 1995. "`Structure of Serum Response Factor Core Bound to DNA."'
\textsl{Nature} \textbf{376:} 490-498.

\bibitem{Prusinkiewicz90}
Prusinkiewicz, P.\ 1990. \textsl{The Algorithmic Beauty of Plants.} Springer Verlag,
New York.

\bibitem{Prusinkiewicz94}
Prusinkiewicz, P. 1994. "`Visual Models of Morphogenesis."' \textsl{Artificial Life}
\textbf{1:} 61-74.

\bibitem{Purugganan95}
Purugganan, M.D., Rounsley, S.D., Schmidt, R.J.\ und Yanofsky, M.F. 1995. "`Molecular Evolution of Flower
Development: Diversification of the Plant MADS-Box Regulatory Gene Family."' \textsl{Genetics}
\textbf{140:} 345-356.

\bibitem{Ray92}
Ray, T.S. 1992. "`An Approach to the Synthesis of Life"'
In \textsl{Artificial Life II}, herausgegeben von C.G.\ Langton 
C.\ Taylor, J.D.\ Farmer und S.\ Rasmussen, 371-408. Santa Fe
Institute Studies in the Sciences of Complexity, Proceedings Volume X.
Addison-Wesley, Redwood City, CA.

\bibitem{Ray94}
Ray, T.S. 1994. "`Evolution, Complexity, Entropy and Artificial Reality."'
\textsl{Physica D} \textbf{71:} 239-266.

\bibitem{Rechenberg}
Rechenberg, I. 1973. \textsl{Evolutionsstrategie.} Friedrich Frommann Verlag, Stuttgart
und Bad Cannstatt.

\bibitem{RozenbergSalomaa86}
Rozenberg, G and Salomaa, A. 1986. \textsl{The Book of L.} Springer Verlag, Heidelberg.

\bibitem{Saedler94}
Saedler, H.\ und Thei"sen, G. 1994. "`'On the origin of species': Mythologische und molekularbiologische
Vorstellungen zur Evolution von Mais."' In \textsl{Jahrbuch 1993,} herausgegeben von B.\ Parthier.
Leopoldina (R.\ 3) \textbf{39:} 261-275.

\bibitem{Schmidt77}
Schmidt, V. 1977. \textsl{Digitalelektronisches Praktikum.} Teubner Studienskripten, B.G.\ Teubner, Stuttgart.

% \bibitem{SchneiderWrede94}
% Schneider, G., Schuchhardt, J.\ und Wrede, P, 1994. "`Peptide Design
% \textsl{in Machina}\/: Development of Artificial Mitochondrial
% Protein Precursor Cleavage Sites by Simulated Molecular Evolution."'
% \textsl{Biophys.\ J.} \textbf{68:} 434-447.

\bibitem{Schneider90}
Schneider, T.D.\ und Stephens, M. 1990. "`Sequence Logos: A New Way to Display Consensus Sequences."'
\textsl{Nucleic Acids Research} \textbf{18:} 6097-6100.

\bibitem{Schroedinger44}
Schr"odinger, E. 1944. \textsl{Was ist Leben?} Lizenzausgabe 1987 f"ur die Wissenschaftliche
Buchgesellschaft, Darmstadt.

\bibitem{Schuster94}
Schuster, P. 1994. "`Extended Molecular Evolutionary Biology: Artificial Life
Bridging the Gap Between Chemistry and Biology."' \textsl{Artificial Life}
\textbf{1:} 39-60.

% \bibitem{Schuster88}
% Schuster, P.\ und Swetina, J. 1988. "`Stationary Mutant Distributions and Evolutionary Optimization."'
% \textsl{Bull. Math. Biol.} \textbf{ 50:} 635-660.

\bibitem{Shannon}
Shannon, C.E.\ und Weaver. W. 1949. \textsl{The Mathematical Theory of
Communication.} University of Illinois Press.

\bibitem{Shore95}
Shore, P.\ und Sharrocks, D. 1995. "`The MADS-Box Family of Transcription Factors."'
\textsl{Eur.\ J.\ Biochem.} \textbf{229:} 1-13.

\bibitem{Sims95}
Sims, K. 1995. "`Evolving 3D Morphology and Behavior by Competition."' \textsl{Artificial
Life} \textbf{1:} 353-372.

\bibitem{Sinha95}
Sinha, N.R., Williams, R.E.\ und Hake, S. 1995. "`Overexpression of the Maize Homeo Box
Gene, KNOTTED-1, Causes a Switch From Determinate to Indeterminate Cell Fates."'
\textsl{Genes \& Development} \textbf{7:} 787-795.

\bibitem{Speicher89}
Speicher, M.\ und Stueber, K. 1989. "`Zur strukturellen Komplexit"at von DNS-Sequenzen."'
Forschungsbericht Nr.\ 320, Fachbereich Informatik der Universit"at Dortmund.

\bibitem{Stanley73}
Stanley, S.M. 1973. "`An Ecological Theory for the Sudden Origin
of Multicellular Life in the Late Precambrian."' \textsl{Proc. Natl. Acad. Sci} 70: 1486-1489.

\bibitem{Taylor94}
Taylor, C.\ und Jefferson, D. 1994. "`Artificial Life as a Tool for Biological Inquiry."'
\textsl{Artificial Life} \textbf{1:} 1-13.

\bibitem{Terzopoulos95}
Terzopoulos, D., Tu, X.\ und Grzeszczuk, R. 1995. "`Artificial Fishes: Autonomous
Locomotion, Perception, Behavior, and Learning  in a Simulated Physical World."'
\textsl{Artificial Life} \textbf{1:} 327-351.

\bibitem{Thearling94}
Thearling, K. 1994. "`Evolution, Entropy, and Parallel computation."' Beitrag zum \textsl{Physics and
Computation Workshop (PhysComp94)}, erh"altlich "uber \textsl{ftp} vom Server \verb|life.slhs.udel.edu|.

\bibitem{Theissen95}
Thei"sen, G.\ und Saedler, H. 1995. "`MADS-box genes in ontogeny and phylogeny
of plants: Haeckel's 'biogenetic law' revisited."' \textsl{Current Opinion in
Genetics and Development} \textbf{5}, im Druck.

\bibitem{Theissen95a}
Thei"sen, G., Kim, J.T\ und Saedler, H. 1995. "`Classification and Phylogeny Reconstruction of
the MADS-Box Multigene Family Suggest Defined Roles of MADS-Box Gene Subfamilies in the Morphological Evolution
of Eukaryotes."' Eingereicht bei \textsl{J.\ Mol.\ Evol.}

\bibitem{Tragut89}
Tragut, V. 1989. "`Graphtale Pflanzen. Simulation von Pflanzenwachstum mit Lindenmeyer-Systemen."'
\textsl{c't} 4/1989: 210-225.

\bibitem{Troebner92}
Tr"obner, W., Ramirez, L., Motte, P., Hue, I., Huijser, P., L"onnig, W.-E., Saedler, H.,
Sommer, H.\ und Schwarz-Sommer, Z. 1992. "`GLOBOSA: A Homeotic Gene which Interacts
with DEFICIENS in the Control of \textsl{Antirrhinum} Floral Organogenesis."'
\textsl{EMBO J.} \textbf{11:} 4693-4704.

\bibitem{Watson}
Watson, J.D., Hopkins, N.H., Roberts, J.W., Argetsinger Steitz, J.\ und Weiner, A.M. 1987. \textsl{Molecular
Biology of the Gene.} Benjamin/Cummings, Menlo Park, CA.

\bibitem{Wilson89}
Wilson, S.W. 1989. "`The Genetic Algorithm and Simulated Evolution."' In \textsl{Artificial Life}, 
herausgegeben von C.G. Langton, 156-166. Santa Fe Institute Studies in the Sciences of 
Complexity, Proceedings Volume VI. Addison-Wesley, Redwood City, CA.


% \bibitem{Ising} {\Large Ising-Modelle...}

% \bibitem{Schreckenberg} {\Large Verkehrsmodell v. Schreckenberg}

\end{thebibliography}

